\documentclass[12pt]{report}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{setspace}
% \usepackage{times}
\usepackage{graphicx}
\graphicspath{ {figures/} }
\usepackage{pgfplots}
\usepackage{url}
\usepackage{xr} % External references
\usepackage{titlesec}

\usepackage{amsmath,amssymb,amsthm,enumitem,mathtools}
\renewcommand{\qedsymbol}{$\blacksquare$}
\usepackage{multirow}
\usepackage{algorithm}% http://ctan.org/pkg/algorithms
\usepackage[noend]{algpseudocode} % Avoid "end" and make it look cleaner

% Tables
\usepackage{csvsimple}	% reading CSV files in tables
\usepackage{booktabs}   % Nicer Tables
\let\oldtabular\tabular
\let\endoldtabular\endtabular
\renewenvironment{tabular}{\renewcommand\arraystretch{1.0}\oldtabular}{\endoldtabular}
\usepackage{adjustbox}  % To adjust table length
\usepackage{float}      % For [H] figure placement
\usepackage{multicol}
\usepackage{verbatim}   % For comment environment
\usepackage{subfig}     % For \subfloat

\usepackage{listings}
\usepackage[cachedir=.]{minted}  % minted v3+ doesn't need finalizecache

% Acronyms
\usepackage[acronym]{glossaries}
\makeglossaries

% Hyperreferences
\usepackage[colorlinks=true, citecolor=cyan, linkcolor=cyan, urlcolor=cyan]{hyperref}

% Equation linebreaks
\allowdisplaybreaks

% Double spacing for main text
\doublespacing

% Single spacing inside multi-line math environments
\usepackage{etoolbox}
% \AtBeginEnvironment{align}{\singlespacing}
% \AtBeginEnvironment{align*}{\singlespacing}
% \AtBeginEnvironment{gather}{\singlespacing}
% \AtBeginEnvironment{gather*}{\singlespacing}

% Custom commands for front matter
\newcommand{\frontmatter}{
    \pagenumbering{gobble} % No page numbers for title and copyright
}
\newcommand{\mainmatter}{
    \pagenumbering{arabic}
    \setcounter{page}{3} % Abstract starts at page 3
}

\input{./definitions.tex}

\begin{document}
\frontmatter

% Title Page (unpaginated)
\begin{titlepage}
\begin{center}
\vspace*{1in}
{\Large \textsc{ Data-Driven Methods for Fast and Reliable Optimization Under Uncertainty}} \\
\vspace{0.5in}
IRINA WANG\\
\vspace{0.5in}
A DISSERTATION \\
PRESENTED TO THE FACULTY \\
OF PRINCETON UNIVERSITY \\
IN CANDIDACY FOR THE DEGREE \\
OF DOCTOR OF PHILOSOPHY\\

\vspace{0.5in}
RECOMMENDED FOR ACCEPTANCE BY\\
THE DEPARTMENT OF \\
\MakeUppercase{Operations Research and Financial Engineering} \\

\vspace{0.5in}
\MakeUppercase{ADVISER: Professor Bartolomeo Stellato}\\
\vspace{0.5in}
MAY 2026 % Use Board of Trustees meeting date
\end{center}
\end{titlepage}

% Copyright Page (unpaginated)
\newpage
\vspace*{1in}
\begin{center}
© Copyright by Irina Wang, 2026. \\ 
All rights reserved.
\end{center}

\newpage

\mainmatter

% Abstract (page 3)
% \chapter*{Abstract}
% \addcontentsline{toc}{chapter}{Abstract}
% Your abstract text here. Must be 350 words or less per Graduate School requirements [1].

\begingroup
\let\clearpage\relax % to avoid starting a new page
\titleformat{\chapter}[block]{\centering\normalfont\Large\bfseries}{\thechapter}{1em}{}
\titlespacing*{\chapter}{0pt}{-4em}{1.5em}

\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}
In today's highly dynamic world, 
% decision-makers must account for changing conditions and unexpected disruptions, lest significant loss be incurred by the lack of robustness.
decision-making under uncertainty has become popular for building solutions that remain effective under various scenarios. However, three main challenges have always limited the practical functionality of proposed methods: \emph{high computational effort} in terms of both memory and runtime, \emph{conservative models} of the uncertainty leading to suboptimal solutions, and \emph{complex formulations} that require advance knowledge of convex analysis.
%  that most practitioners do not possess.
Using tools from optimization and machine learning, this thesis addresses these challenges by designing customized data-driven frameworks.


In the first part of this thesis, we employ data-compression techniques to provide a tradeoff between computational effort and conservatism. 
We construct uncertainty sets 
based on clustered data, and by varying the number of clusters, our
method bridges between \gls{RO} and Wasserstein \gls{DRO}. 
We provide finite-sample probabilistic guarantees of constraint satisfaction, and derive bounds on the effect of clustering.
Several numerical examples show that our method provides multiple orders of magnitude speedups in solution time with little-to-no effect on the solution quality. We extend this work to the online setting, where sequential acquisition of data-points and dynamic adaptation of the uncertainty set is considered. We prove asymptotic performance guarantees for solutions under generalized online clustering algorithms.
In the second part of this thesis, we employ end-to-end procedures to learn problem and context-specific uncertainty formulations. Our method reshapes the uncertainty sets by optimizing for both worst-case and average-case performance while guaranteeing constraint satisfaction via a conditional-value-at-risk constraint. We solve the resultant constrained learning problem using a stochastic augmented Lagrangian method.
Numerical experiments show that our method outperforms traditional approaches in \gls{RO} and \gls{DRO} in terms of out-of-sample performance and constraint satisfaction guarantees.
In the final part of this thesis, we provide an open-source Python package for applying these frameworks in the real-world, greatly reducing the mathematical overhead associated with optimization under uncertainty.

\newpage
\titlespacing*{\chapter}{0pt}{0em}{1.5em}
\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}
% Throughout my time at Princeton, I had the fortune 
First and foremost, I am deeply grateful to my advisor, Professor Bartolomeo Stellato, for his unwavering support and dedicated guidance throughout my Ph.D. journey. 
No matter at what time, his door (or Slack/Zulip account) is always open for discussions on high-level concepts, technical details, coding implementations, presentation slides, as well as career related advice. 
He has been an incredible mentor and a constant source of inspiration, and has, on countless occasions, pointed me to opportunities I never knew existed. 
His positivity and enthusiasm have not only provided me with motivation for our projects, but have also fostered a healthy mindset with which I can face any challenge. 
This dissertation and my future career would not have been the same without him. 


I am also incredibly grateful to Professor Bart P. G. Van Parys, for his mentoring and support throughout the years, and particularly for his willingness to join zoom meetings on short notice to provide valuable insights. 
I am especially thankful to Professor Daniel Kuhn for hosting me for a summer at EPFL. The visit to his research group was an enlightening and enriching experience, both for the research insighted gained and for the refreshing change of pace that life in Switzerland provided. 


My time in the ORFE department has been made enjoyable by the brilliant and kind faculty members and staff. I would like to thank my Ph.D. committee members Professor Jason Klusowski and Professor Amir Ali Ahmadi, and my dissertation reader Professor Elizaveta Rebrova, for their valuable time and insightful remarks not only on my thesis, but also for my practice job talk in the prior months. My deep appreciation also extends to the administrative staff, who have made my life significanly easier. 
In particular, I would like to thank Kimberly Lupinacci for keeping things running and accommodating my random requests, and Michael Bino for the magic he performs to solve any and all OIT related issues.

It has been a great pleasure to be a part of Bartolomeo's research group. To Rajiv Sambharya, Vinit Ranjan, and Stefan Clarke: I'll always remember the early years, when we were all new and relatively inexperienced, travelling together for our first conferences and sharing various AirBnBs. It has been rewarding to see each other's growth amidst our mutual encouragement and deepening friendships. Thank you also to the other members and affiliates of the group, who have filled my days with warmth and laughter: Yixuan Hua, Yanjun Liu, Daniel Deza, Jisun Park, Aras Selvi, Gabriele Dragotto, Cole Becker, Liubomir Baicev, Amit Solomon, and Paul Häusner. I have learned a lot from all of you through our plentiful interactions, whether in our group meetings, office hangouts, or beyond. 


I am grateful to my first office mate Suqi Liu, who really showed me the ropes when I arrived at Princeton, and made me feel welcome to the new community. I would also like to thank my subsequent office mates, Sofiia Shvaiko, Giang Truang, and Erica Lai, who made our ``all-girls" office a truly wonderful experience. 
My time at Princeton was made all the richer by my friends and fellow graduate students in the ORFE department: Abraar Chaudhry, Kaiwen Zhang, Benjamin Budway, Carla Crucianelli,  Nicolas Garcia, Felix Hoefer, Giulia Crippa, Kevin Wong, Yukai Tang, Jiawei Ge, Silvia Tao, and many others, who have made Sherrerd Hall such a lively and inspiring space to be.


Last but not least, I would like to thank my partner, long-time friends, and family, for their unconditional and steadfast support. Their faith in me has given me incredible strength throughout this long journey. 

To my parents, Baoping Wang and Jun Zheng: I would not be who I am today without you.



\endgroup

\newpage

% Table of Contents
\tableofcontents

% List of Figures (if needed)
\listoffigures
\addcontentsline{toc}{chapter}{List of Figures}

% List of Tables (if needed)
\listoftables
\addcontentsline{toc}{chapter}{List of Tables}

% Set array stretch for mathematical content only
\renewcommand\arraystretch{1.2}

% Main content starts here
% Page numbering continues from front matter

\include{chapter1}
\include{chapter2}
\include{chapter3}
\include{chapter4}
\include{chapter5}
\include{chapter6}
\include{chapter7}
% Bibliography
\bibliographystyle{abbrv}
% \bibliographystyle{natbib}
\bibliography{references}
\addcontentsline{toc}{chapter}{Bibliography}

% Appendices (if needed)
\appendix
\include{appendix}

\end{document}
